An equivalence of compositional frameworks

### 2019-05-02 EquivCompFrameworks_20180911kc.tex
wrote down all pieces of BRIDGE!

\begin{thm}

@@

\section{An equivalence of symmetric monoidal double categories} \label{EquivDoubleCats}
In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and structured cospans described in \cref{sec:structuredcospans,DecCospansDoubleCat}, are equivalent. This is established by \cref{Equiv}, under some assumptions relevant to the existence of left adjoints for opfibrations. We begin by elaborating this underlying framework, using results mainly from \cite{CV} and \cite{MV}.

@@

\begin{thm}\label{thm:Grothendieck}\hfill
    \begin{enumerate}
        \item Every opfibration $\U \maps \D \to \C$ gives rise to a pseudofunctor $F_U \maps \C {\to}Cat$.
        \item Every opindexed category $F \maps \C \to Cat$ gives rise to  an opfibration $\U_F \colon \int F\to\C$.
        \item The above correspondences yield an equivalence of 2-categories 
        \begin{displaymath}
            ICat(\C) \simeq opFib(\C)
        \end{displaymath}
        so that $F_{\U_F} \cong F$ and $\U_{F_\U} \cong \U$.
%        \item The above 2-equivalence extends to one between 2-categories of arbitrary-base fibrations and arbitrary-domain indexed categories
%        \begin{equation}\label{eq:Gr_equiv}
%        \ICat\simeq\Fib    
%        \end{equation}
    \end{enumerate}
\end{thm}

@@

{\chris A little bit for monoidal Grothendieck here}

For our purposes, we are interested in the fact that under certain conditions, the induced monoidal opfibration by the Grothendieck construction $ \int{F} \to \A$ will be a right adjoint to some adjoint $L_F \colon \A \to \X$, and moreover the unit of the adjunction is the identity; this is called \emph{lari} in [Gray].

{\chris in progress, probably split up in two}
\begin{thm}\label{thm:mainthmDX}\cite{CV}
Suppose that $\X$ and $\A$ have chosen pushouts and initial objects, and a functor $\U\colon\X\to\A$ strictly preserves them. Then $\U$ is a right adjoint left inverse if and only if $\U$ is an opfibration.
\end{thm}

In the above described setting, a symmetric lax monoidal functor gives rise to a double category of decorated cospans as well as a double category of structured cospans. The result below establishes their equivalence.

\begin{thm}\label{Equiv}
Let $\mathsf{A}$ and $\mathsf{X}$ be categories with finite colimits, and $F \colon(\mathsf{A},+,0) \to (\mathbf{Cat},\times,\mathbf{1})$ a symmetric lax monoidal pseudofunctor. The symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of decorated cospans is equivalent to the symmetric monoidal double category $_{L_F} \mathbb{C}\textnormal{sp}(\mathsf{X})$ of structured cospans, where $L_F \colon \mathsf{A} \to \int{F}$ is the left adjoint of the induced Grothendieck opfibration $\U_F\colon \int{F} \to \mathsf{A}$ by \cref{thm:Grothendieck} under the assumptions of \cref{thm:mainthmDX}.
\end{thm}
\begin{proof}
As explained in (ref), the double category of decorated cospans $\dcsp{F}$ has $\A$ as its category of objects and vertical 1-morphisms, whereas horizontal 1-cells $a\tickar b$ are cospans $\cspn{a}{c}{b}{}{}$ in $\A$ along with $d\in F(c)$ and 2-morphisms are cospan maps $k\colon c\to c'$ such that $(Fk)(d)=d'$.

@@

{\chris in progress}
To prove this, we define a double functor $\mathbb{E} \colon _L \mathbb{C}\textnormal{sp}(\mathsf{X}) \to F\mathbb{C}\textnormal{sp}$ as follows: the object component of the double functor $\mathbb{E}$ is given by $\mathbb{E}_0 = \id_{\mathsf{A}}$ as both double categories $_L \mathbb{C}\textnormal{sp}(\mathrm{X})$ and $F\mathbb{C}\textnormal{sp}$ have objects and morphisms of $\mathsf{A}$ as objects and vertical 1-morphisms, respectively. The functor $\mathbb{E}_0$ is trivially an equivalence of categories.

@@

@@

@@

>>>>

\begin{thm}\label{thm:decoratedcospans}

@@

\section{An equivalence of symmetric monoidal double categories} \label{EquivDoubleCats}
In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and structured cospans described in \cref{sec:structuredcospans,DecCospansDoubleCat}, are equivalent under certain assumptions. This is established by \cref{Equiv} which is based on a framework relevant to the existence of left adjoints for opfibrations. We begin by sketching this underlying framework, which is worked out in detail in \cite{CV} and \cite{MV}. Standard material on fibrations and the Grothendieck construction can be found for example in \cite{Herm}, \cite{Borc}.

@@

\begin{thm}\label{thm:Grothendieck}\hfill
    \begin{enumerate}
        \item Every opfibration $\U \maps \D \to \C$ gives rise to a pseudofunctor $F_U \maps \C \to \bicat{Cat}$.
        \item Every opindexed category $F \maps \C \to Cat$ gives rise to  an opfibration $\U_F \colon \int F\to\C$.
        \item The above correspondences yield an equivalence of 2-categories 
        \begin{displaymath}
            ICat(\C) \simeq opFib(\C)
        \end{displaymath}
        so that $F_{\U_F} \cong F$ and $\U_{F_\U} \cong \U$.
%        \item The above 2-equivalence extends to one between 2-categories of arbitrary-base fibrations and arbitrary-domain indexed categories
%        \begin{equation}\label{eq:Gr_equiv}
%        \ICat\simeq\Fib    
%        \end{equation}
    \end{enumerate}
\end{thm}

@@

In \cite{MV}, the Grothendieck construction is generalized to the monoidal setting, namely lax monoidal structures on a pseudofunctor $F\colon\A\to\bicat{Cat}$ bijectively correspond to monoidal structures on the total category $\int F$ such that the induced opfibration is a strict monoidal functor and $\otimes_{\int F}$ is cartesian, see \cite[Thm. 3.10]{MV}. If the monoidal category $\A$ is in fact cocartesian monoidal, there is a further correspondence of the two structures with a sheer pseudofunctor $F\colon\A\to\ca{MonCat}$:
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\colon(\A,+,0)\to(\bicat{Cat},\times,1) \\
\Updownarrow \\
\textrm{monoidal opfibrations }U\colon(\X,\otimes,I)\to(\A,+,0) \label{monGroth}\\
\Updownarrow \\
\textrm{pseudofunctors }F\colon\A\to\bicat{MonCat}
\end{gather}
where the second equivalence was already observed by Shulman in \cite{Shul2}. Moreover, the above further restricts to the case when the Grothendieck category is specifically cocartesian monoidal itself, see \cite[Cor. 4.8]{MV} for more details. Then, the monoidal opfibration clauses for $U\colon(\X,+,0)\to(\A,+,0)$ in fact translate to the functor (strictly) preserving coproducts and the initial object, and bijectively correspond to pseudofunctors $F\colon\A\to\bicat{cocartCat}$ into the 2-category of cocartesian categories, functors that preserve coproducts and initial objects and natural transformations.

Finally, using the following result we can further restrict to our current case of interest, namely opfibrations that also preserve pushouts, i.e. all finite colimits. Of course this is more general since it relates the existence of any class of colimits in the total category of an opfibration to their existence in the fibres; for more details, see \cite{Herm}.

\begin{lem}\label{lem:fibrewiselimits}
Suppose $\ca{J}$ is a small category and $\U\colon\X\to\A$ is an opfibration. If $\A$ has $\ca{J}$-colimits,
the following are equivalent:
\begin{enumerate}
 \item all fibres have $\ca{J}$-colimits, and the reindexing functors preserve them;
 \item $\X$ has $\ca{J}$-colimits, and $\U$ strictly preserves them.
\end{enumerate}
\end{lem}

The first part formulates the existence of colimits \emph{locally} in each fiber, which can equivalently be expressed by the corresponding pseudofunctor $F\colon\A\to\bicat{Cat}$ landing on the sub-2-category $\bicat{fcocCat}$ of finitely cocomplete categories and functors that preserve colimits; the second part formulates the existence of colimits \emph{globally} in the total category $\int F$. Combining these, we can deduce the following.

\begin{cor}\label{cor:fcocMonGroth}
 Suppose $\A$ is a finitely cocomplete category and $F\colon(\A,+,0)\to(\bicat{Cat},\times,1)$ is a lax monoidal pseudofunctor. If its corresponding pseudofunctor $\A\to\bicat{MonCat}$ via \cref{monGroth} 
 factors through $\bicat{fcocCat}$, then the Grothendieck category $\int F$ has all finite colimits and the induced opfibration $\U_F\colon\int F\to\A$ preserves them. 
\end{cor}

On a different but related subject, for our purposes we are interested in the existence of a left adjoint $L_F \colon \A \to \X$ for the induced monoidal opfibration $ \int{F} \to \A$ by the Grothendieck construction as discussed above. The following result provides sufficient conditions for that, which go hand-in-hand with \cref{cor:fcocMonGroth}.

\begin{prop}\cite[Prop.\ 4.4]{Gray} \label{prop:opfibtolari}
  Let $\U \colon\D \to \C$ be an opfibration. Then
  $\U$ is a right adjoint `left inverse', a.k.a the unit is the identity, if (and only if) its fibers have
  initial objects which are preserved by the
  reindexing functors.
\end{prop}
%will be a right adjoint to some adjoint , and moreover the unit of the adjunction is the identity; this is called \emph{lari} in [Gray].

In the opposite direction, we also have the following result. For a discussion on the unusual strict cocontinuity condition, we refer the reader to \cite{CV}.

\begin{prop}\label{prop:CV}\cite[Prop. \ 3.3]{CV}
Suppose $\U\colon\X\to\A$ is a right adjoint `left inverse'. If $\X$ and $\A$ have chosen pushouts and initial objects and $\U$ strictly preserves them, then $\U$ is an opfibration.
\end{prop}
%Christina: do we need this direction after all? Discuss.

%\begin{thm}\label{thm:mainthmDX}\cite{CV}
%Suppose that $\X$ and $\A$ have chosen pushouts and initial objects, and a functor $\U\colon\X\to\A$ strictly preserves them. Then $\U$ is a right %adjoint left inverse if and only if $\U$ is an opfibration.
%\end{thm}

We have now laid all the necessary theoretical background to formally establish an equivalence between the double category of decorated cospans and the double category of structured cospans constructed from a symmetric lax monoidal pseudofunctor into $\bicat{Cat}$.

\begin{thm}\label{Equiv}
Suppose $\mathsf{A}$ and $\mathsf{X}$ are finitely cocomplete categories and $F \colon(\mathsf{A},+,0) \to (\mathbf{Cat},\times,\mathbf{1})$ is a symmetric lax monoidal pseudofunctor. %such that as an ordinary pseudofunctor, it factors through $\bicat{fcocCat}$. 
The symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of decorated cospans (\cref{thm:decoratedcospans}) is equivalent to the symmetric monoidal double category $_{L_F} \mathbb{C}\textnormal{sp}(\mathsf{X})$ of structured cospans (\cref{SC}), where $L_F \colon \mathsf{A} \to \int{F}$ is the left adjoint of the induced Grothendieck opfibration $\U_F\colon \int{F} \to \mathsf{A}$ by \cref{thm:Grothendieck}, under the assumptions of \cref{cor:fcocMonGroth}.
\end{thm}
\begin{proof}
{\chris in progress}
As explained in (ref), the double category of decorated cospans $\dcsp{F}$ has $\A$ as its category of objects and vertical 1-morphisms, whereas horizontal 1-cells $a\tickar b$ are cospans $\cspn{a}{c}{b}{}{}$ in $\A$ along with $d\in F(c)$ and 2-morphisms are cospan maps $k\colon c\to c'$ such that $(Fk)(d)=d'$.

@@

To prove this, we define a double functor $\mathbb{E} \colon _L \mathbb{C}\textnormal{sp}(\mathsf{X}) \to F\mathbb{C}\textnormal{sp}$ as follows: the object component of the double functor $\mathbb{E}$ is given by $\mathbb{E}_0 = \id_{\mathsf{A}}$ as both double categories $_L \mathbb{C}\textnormal{sp}(\mathrm{X})$ and $F\mathbb{C}\textnormal{sp}$ have objects and morphisms of $\mathsf{A}$ as objects and vertical 1-morphisms, respectively. The functor $\mathbb{E}_0$ is trivially an equivalence of categories.

@@

\bibitem{Borc} F. \ Borceux, Handbook of categorical algebra. 2, volume 51 of Encyclopedia of Mathematics and its Applications,
Cambridge University Press, 1994

@@

\bibitem{Gray} J. \ Gray, Fibred and Cofibred categories, Proc. Conf. Categorical Algebra (La Jolla, California, 1965), 21---83, New York, 1966

@@

\bibitem{Herm} C. \ Hermida, Fibrations, logical predicates and indeterminates, PhD Thesis, University of Edinburgh, 1993.
